---
title: "Week 2 In-Class Exercise: Vectors & Subsetting"
subtitle: "GPH-GU 2182 · Statistical Programming in R"
author: "YOUR NAME & PARTNER NAME"
date: "September 9, 2026"
format:
  html:
    toc: true
    number-sections: true
    code-tools:
      source: true
      toggle: false
      caption: "Source Code"
---

::: {.callout-note}
### Instructions
Work in pairs during class. 

* 📥 **[Download `week-02-inclass.qmd`](https://raw.githubusercontent.com/gph-2182/gph-2182.github.io/main/weeks/inclass/week-02-inclass.qmd)** to open and edit in RStudio (or use the **`</> Source Code`** button at the top right).
* Render this Quarto document to HTML and push it to your GitHub repository by the end of class for participation credit.
:::

# Cycle 1: Vector Types & Implicit Coercion

## Type Predictions and Verification

In the code chunk below, write R code to check the `typeof()` of each of the following expressions:

- `vec_a <- c(4, 7, FALSE)`
- `vec_b <- c(3L, 5.5, "control")`
- `vec_c <- c(TRUE, FALSE) + 5`

```{r type-checks}
# Create the vectors and check their types:

```

## Public Health Prevalence Calculation

A study recorded the HIV viral load copies/mL for 8 participants:
`1200, 450, 0, 8900, 320, 0, 1500, 60`.

1. Create a numeric vector named `viral_load` with these measurements.
2. Use a logical comparison and the `mean()` function to calculate the **proportion of patients with undetectable viral load** ($\text{viral load} = 0$).

```{r viral-load}
# Calculate undetectable proportion:

```

---

# Cycle 2: Creating Patterned Vectors

## Dosage Schedules with `seq()`

An intravenous antibiotic is administered every **8 hours** across a **72-hour** hospital observation period.

Create a sequence named `dose_hours` starting at hour 0 up to hour 72 with increments of 8.

```{r dose-seq}
# Create dosage hours:

```

## Cohort Generation with `rep()`

A vaccine trial requires creating treatment labels for 12 participants:
- 4 "Control"
- 4 "Low-Dose"
- 4 "High-Dose"

Generate this character vector using the `rep()` function and assign it to `treatment_groups`.

```{r treatment-rep}
# Generate treatment labels:

```

## Extracting Multiples with Modulo `%%`

Given the sequence `numbers <- 1:50`, extract all numbers that are divisible by 7.

```{r modulo-filter}
numbers <- 1:50
# Extract multiples of 7:

```

---

# Cycle 3: Subsetting & Mutating Vectors

## Glucose Filtering and Cleaning

A metabolic health study collected fasting blood glucose levels (mg/dL) from 8 patients:
`glucose <- c(95, 142, 180, 88, 210, 115, -1, 104)`

1. Extract all values in the **pre-diabetes range** ($100 \le \text{glucose} \le 125$).
2. The measurement `-1` is a missing value code. Replace all negative values in `glucose` with `NA`.
3. Extract all valid readings ($\ge 126$, excluding `NA`) indicating **diabetes**.

```{r glucose-subset}
glucose <- c(95, 142, 180, 88, 210, 115, -1, 104)

# 1. Pre-diabetes subset:

# 2. In-place replacement of negative values with NA:

# 3. Diabetes subset (>= 126, non-NA):

```

---

# Cycle 4: Sorting, Ranking & AI Verification

## Patient Rankings

A clinic recorded wait times (minutes) for 5 emergency room patients:
`wait_times <- c(45, 12, 85, 30, 45)`

1. Sort the wait times in **descending order** (longest to shortest).
2. Compute the ranks of each patient's wait time using `rank(..., ties.method = "first")`.

```{r wait-times}
wait_times <- c(45, 12, 85, 30, 45)

# 1. Descending sort:

# 2. First-appearance ranking:

```

## AI Corner: Spotting Vector Recycling Errors

An AI assistant was asked to calculate the Body Mass Index for 4 patients with heights (in meters) `c(1.70, 1.85, 1.62, 1.78)` and weights (in kg) `c(65, 82)`.

The AI produced this code:

```{r ai-recycling}
# AI Generated code:
heights <- c(1.70, 1.85, 1.62, 1.78)
weights <- c(65, 82)
bmi_calc <- weights / (heights^2)
```

### Questions:

1. What did R do when computing `bmi_calc`? Why is the result clinically invalid?
2. How would you write a defensive check in R to prevent vector recycling when calculating BMI?

*(Write your explanation and defensive check below)*

---

## Submission Checklist

- [ ] Both your name and your partner's name are in the `author` field at the top.
- [ ] Click **Render** in RStudio and verify the HTML compiles cleanly.
- [ ] Commit and push your changes to your GitHub repository.
