---
title: "Week 5 In-Class Exercises: Data Wrangling with dplyr & Pipes"
subtitle: "GPH-GU 2182 · Fall 2026 · Paired Programming"
author: "YOUR NAME & PARTNER NAME"
date: "September 30, 2026"
format:
  html:
    toc: true
    number-sections: true
    code-tools:
      source: true
      toggle: false
      caption: "Source Code"
---

::: {.callout-note}
### Instructions
Work in pairs. One person types (driver), the other reviews and advises (navigator). Swap roles halfway through!

* 📥 **[Download `week-05-inclass.qmd`](https://gph-2182.github.io/weeks/inclass/week-05-inclass.qmd)** to open and edit in RStudio (or use the **`</> Source Code`** button at the top right).
* Render this Quarto document to HTML and push it to your GitHub repository by the end of class for participation credit.
:::

```{r}
#| label: setup
#| message: false
library(r02pro)
library(dplyr)
```

# Exercise 1: High Male Smoking Rates (Cycle 1)

Use the `gm` global health dataset to find the countries with the heaviest male smoking burden in 2008.

```{r}
#| label: ex1-filter-arrange
# 1. Starting from gm, filter to rows for the year 2008
# 2. Keep only rows where smoking_male is not missing (!is.na())
# 3. Keep only rows where smoking_male > 50
# 4. Arrange from highest to lowest smoking_male
# 5. Report the number of remaining countries with nrow()

```

# Exercise 2: A Tidy Lung Cancer Table (Cycle 2)

Build a compact table of lung cancer incidence by country.

```{r}
#| label: ex2-select-rename
# 1. Filter gm to the year 2008
# 2. Select country, continent, and every column that contains "lungcancer"
# 3. Rename lungcancer_newcases_female to lung_f and lungcancer_newcases_male to lung_m
# 4. Arrange by lung_m in descending order and display the top rows with head()

```

# Exercise 3: Food Supply & Longevity (Cycle 3)

Investigate how food supply relates to life expectancy using new derived variables.

```{r}
#| label: ex3-mutate
# 1. Filter gm to 2008 rows with non-missing food_supply and life_expectancy
# 2. Select country, continent, food_supply, life_expectancy
# 3. Use one mutate() call to create BOTH:
#      food_kcal_deficit = 2500 - food_supply
#      long_lived        = life_expectancy >= 75
# 4. Arrange by food_kcal_deficit (largest deficit first) and inspect the top 10 rows

```

# Exercise 4: Full Pipeline + AI Corner (Cycle 4)

An AI assistant claims the following pipeline computes *"sugar consumption per continent"*, but it uses tools we have not verified and contains bugs. Rebuild it correctly using only Week 1–5 tools.

```{r}
#| label: ex4-pipeline
# The AI's (flawed) suggestion:
#   gm %>% group_by(continent) %>% summarize(mean_sugar = mean(sugar))
#
# Your corrected, verifiable version:
# 1. Filter gm to year == 2008, continent == "Europe", and non-missing sugar
# 2. Select country and sugar; rename sugar to sugar_g_per_day
# 3. Assign the pipeline result to europe_sugar
# 4. Compute mean(europe_sugar$sugar_g_per_day) and compare with a classmate:
#    can you explain every line of your pipeline?

```
