---
title: "Week 4 In-Class Exercises: Data Visualization with ggplot2"
subtitle: "GPH-GU 2182 · Fall 2026 · Paired Programming"
format:
  html:
    toc: true
    number-sections: false
---

::: {.callout-note}
### Instructions
Work in pairs. One person types (driver), the other reviews and advises (navigator). Swap roles halfway through! Push your final rendered `.qmd` and `.html` to your GitHub repository by the end of class for participation credit.
:::

<a class="btn btn-primary mb-3" href="https://raw.githubusercontent.com/gph-2182/course-website/main/weeks/inclass/week-04-inclass.qmd" download="week-04-inclass.qmd" role="button">
📥 Download Starter Code (`.qmd`)
</a>

```{r}
#| label: setup
#| message: false
library(ggplot2)
```

---

## Exercise 1: Environmental Air Quality Scatterplot (Cycle 1)

Using the built-in `airquality` dataset, visualize the relationship between solar radiation and ozone concentrations.

```{r}
#| label: ex1-scatterplot
# 1. Create a scatterplot of Solar.R (x) vs Ozone (y)
# 2. Map Wind speed to the point color aesthetic
# 3. Set alpha = 0.7 and size = 3 outside aes()
# 4. Add a smooth trendline with geom_smooth(method = "loess", color = "darkblue")

ggplot(airquality, aes(x = Solar.R, y = Ozone)) +
  geom_point(aes(color = Wind), size = 3, alpha = 0.7) +
  geom_smooth(method = "loess", color = "darkblue", se = TRUE) +
  labs(
    title = "Ozone Pollution vs. Solar Radiation",
    x = "Solar Radiation (Langleys)",
    y = "Ozone Concentration (ppb)",
    color = "Wind (mph)"
  ) +
  theme_minimal()
```

---

## Exercise 2: Comparing Ozone Across Summer Months (Cycle 2)

Compare the distribution of ozone levels across the five summer months (May = 5 through September = 9).

```{r}
#| label: ex2-distributions
# 1. Build a violin plot comparing Ozone across Months (x = factor(Month), y = Ozone)
# 2. Fill the violins by factor(Month) with alpha = 0.5
# 3. Overlay a boxplot inside the violins with width = 0.2 and fill = "white"
# 4. Suppress duplicate outlier points on the boxplot using outlier.shape = NA

ggplot(airquality, aes(x = factor(Month), y = Ozone, fill = factor(Month))) +
  geom_violin(alpha = 0.5, trim = FALSE) +
  geom_boxplot(width = 0.2, fill = "white", outlier.shape = NA, alpha = 0.8) +
  scale_x_discrete(labels = c("May", "Jun", "Jul", "Aug", "Sep")) +
  labs(
    title = "Monthly Ozone Distribution in NYC (1973)",
    x = "Month",
    y = "Ozone (ppb)",
    fill = "Month"
  ) +
  theme_minimal()
```

---

## Exercise 3: Small Multiples with `facet_wrap` (Cycle 3)

Examine how the temperature-ozone relationship changes month by month.

```{r}
#| label: ex3-facets
# 1. Create a scatterplot of Temp (x) vs Ozone (y)
# 2. Add points with alpha = 0.6
# 3. Add a linear regression line using geom_smooth(method = "lm", color = "red", se = FALSE)
# 4. Facet by Month into a single row using facet_wrap(~ Month, nrow = 1)

ggplot(airquality, aes(x = Temp, y = Ozone)) +
  geom_point(alpha = 0.6, color = "navy") +
  geom_smooth(method = "lm", color = "red", se = FALSE) +
  facet_wrap(~ Month, nrow = 1) +
  labs(
    title = "Temperature vs. Ozone Concentration Across Months",
    x = "Max Daily Temperature (°F)",
    y = "Ozone (ppb)"
  ) +
  theme_bw()
```

---

## Exercise 4: Publication-Ready Polishing (Cycle 4 + AI Corner)

Take the multi-panel temperature plot from Exercise 3 and polish it to publication standards.

```{r}
#| label: ex4-publication
# 1. Map Wind speed to color using scale_color_viridis_c(option = "viridis")
# 2. Add complete titles, subtitle, axis labels, and caption
# 3. Apply theme_minimal() and position the legend at the bottom

p_pub <- ggplot(airquality, aes(x = Temp, y = Ozone, color = Wind)) +
  geom_point(size = 2.5, alpha = 0.8) +
  geom_smooth(method = "lm", color = "black", linetype = "dashed", se = FALSE) +
  scale_color_viridis_c(option = "viridis") +
  facet_wrap(~ Month, nrow = 1) +
  labs(
    title = "Air Pollution Dynamics at LaGuardia Airport",
    subtitle = "Daily measurements of Ozone, Temperature, and Wind (May - Sep 1973)",
    x = "Temperature (°F)",
    y = "Ozone (ppb)",
    color = "Wind (mph)",
    caption = "Source: NYS Department of Environmental Conservation"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "bottom",
    plot.title = element_text(face = "bold", size = 14)
  )

p_pub
```
