---
title: "Week 8 In-Class Exercises: Control Flow & Iteration"
subtitle: "GPH-GU 2182 · Fall 2026 · Paired Programming"
author: "YOUR NAME & PARTNER NAME"
date: "October 28, 2026"
format:
  html:
    toc: true
    number-sections: true
    code-tools:
      source: true
      toggle: false
      caption: "Source Code"
---

::: {.callout-note}
### Instructions
Work in pairs. One person types (driver), the other reviews and advises (navigator). Swap roles halfway through!

* 📥 **[Download `week-08-inclass.qmd`](https://raw.githubusercontent.com/gph-2182/gph-2182.github.io/main/weeks/inclass/week-08-inclass.qmd)** to open and edit in RStudio (or use the **`</> Source Code`** button at the top right).
* Render this Quarto document to HTML and push it to your GitHub repository by the end of class for participation credit.
:::

```{r}
#| label: setup
#| message: false
library(dplyr)
```

# Exercise 1: Blood Pressure Stage Classifier (Cycle 1)

```{r}
#| label: ex1-data
sbp_clinic <- c(118, 152, 141, 129, 183, 136, 147)
```

```{r}
#| label: ex1-ifelse
# 1. Assign sbp <- 152 and write an if / else if / else chain that prints:
#    "crisis" if sbp >= 180, "stage 2" if >= 140, "stage 1" if >= 130, else "normal"
# 2. Re-run your chain with sbp <- 128 and sbp <- 185 to confirm each branch
# 3. For the sbp_clinic vector, use ifelse() to label each reading
#    "refer" (>= 140) or "routine"
# 4. Rewrite step 3 with case_when() adding a third tier: "crisis" (>= 180)

```

# Exercise 2: Outbreak Accumulator (Cycle 2)

```{r}
#| label: ex2-data
daily_cases <- c(3, 7, 12, 20, 31, 26, 18, 40, 22, 15)
```

```{r}
#| label: ex2-forloop
# 1. Write a for loop with an accumulator that builds cumulative_cases
#    (pre-allocate with numeric(length(daily_cases)), index with seq_along())
# 2. Verify your result equals cumsum(daily_cases) using identical() or all.equal()
# 3. Inside the SAME loop, maintain a second counter: how many days exceeded 25 cases?
# 4. Verify the counter equals sum(daily_cases > 25)

```

# Exercise 3: Elimination Countdown with while (Cycle 3)

A vaccination campaign reduces monthly measles cases by 35%, starting from 5,000 cases.

```{r}
#| label: ex3-while
# 1. Use a while loop to count how many months until cases fall below 50
# 2. Record the case count history in a vector as the loop runs
# 3. Add a guard rail so the loop can never run more than 100 months
# 4. Re-run with a 10% monthly reduction: how many months now?

```

# Exercise 4: Looping Over Groups + AI Corner (Cycle 4)

```{r}
#| label: ex4-data
#| message: false
library(r02pro)
```

```{r}
#| label: ex4-integration
# 1. Write a for loop over continents <- c("Africa", "Asia", "Europe") that, for each,
#    filters gm to year 2008, that continent, non-missing life_expectancy >= 75,
#    and prints "<continent> has <n> countries with life expectancy >= 75"
# 2. AI Corner: an AI wrote this to flag high-mortality countries:
#      flags <- c()
#      for (i in 1:nrow(df)) { flags <- c(flags, if (df$child_mortality[i] > 50) "high" else "low") }
#    List two problems with it, then write the one-line vectorized replacement
#    using ifelse() on gm's child_mortality for 2008.

```
