---
title: "Week 6 In-Class Exercises: Advanced Wrangling, Joins & Tidy Data"
subtitle: "GPH-GU 2182 · Fall 2026 · Paired Programming"
author: "YOUR NAME & PARTNER NAME"
date: "October 7, 2026"
format:
  html:
    toc: true
    number-sections: true
    code-tools:
      source: true
      toggle: false
      caption: "Source Code"
---

::: {.callout-note}
### Instructions
Work in pairs. One person types (driver), the other reviews and advises (navigator). Swap roles halfway through!

* 📥 **[Download `week-06-inclass.qmd`](https://raw.githubusercontent.com/gph-2182/gph-2182.github.io/main/weeks/inclass/week-06-inclass.qmd)** to open and edit in RStudio (or use the **`</> Source Code`** button at the top right).
* Render this Quarto document to HTML and push it to your GitHub repository by the end of class for participation credit.
:::

```{r}
#| label: setup
#| message: false
library(r02pro)
library(dplyr)
library(tidyr)
library(tibble)
```

# Exercise 1: Continental Report Card (Cycle 1)

Summarize the global burden of child mortality by continent using `gm`.

```{r}
#| label: ex1-group-summarize
# 1. From gm, keep 2008 rows with non-missing child_mortality
# 2. Group by continent
# 3. Summarize: mean_cm = mean(child_mortality), max_cm = max(child_mortality),
#    n_countries = n()
# 4. Arrange the result from highest to lowest mean_cm

```

# Exercise 2: Lab Results Lookup with Joins (Cycle 2)

A clinic network stores visits and lab results in separate tables.

```{r}
#| label: ex2-data
visits <- tibble(
  visit_id = c("V01", "V02", "V03", "V04", "V05", "V06"),
  clinic   = c("Downtown", "Downtown", "Uptown", "Uptown", "Midtown", "Midtown"),
  age      = c(52, 34, 68, 45, 59, 71)
)

lab_results <- tibble(
  visit_id = c("V01", "V03", "V04", "V06"),
  glucose  = c(112, 145, 98, 160)
)
```

```{r}
#| label: ex2-joins
# 1. left_join() visits and lab_results by "visit_id". How many rows result? Why?
# 2. Use anti_join() to list the visits with NO lab result on file
# 3. Use inner_join(), then group_by(clinic) |> summarize(mean_glucose = mean(glucose))

```

# Exercise 3: Reshaping a Blood Pressure Study (Cycle 3)

A hypertension study measured systolic blood pressure at three follow-up visits (wide format).

```{r}
#| label: ex3-data
bp_wide <- tibble(
  patient   = c("P1", "P2", "P3", "P4"),
  bp_visit1 = c(142, 129, 155, 137),
  bp_visit2 = c(138, 125, 149, 133),
  bp_visit3 = c(131, 122, 144, 128)
)
```

```{r}
#| label: ex3-pivot
# 1. Pivot bp_wide into long form bp_long with columns patient, visit, bp
#    (use cols = starts_with("bp_"), names_prefix = "bp_" to strip the prefix)
# 2. Group bp_long by visit and summarize mean_bp = mean(bp)
# 3. Pivot bp_long back to wide format and confirm it matches bp_wide

```

# Exercise 4: Integration Challenge + AI Corner (Cycle 4)

Which continents improved life expectancy the most between 1998 and 2008?

```{r}
#| label: ex4-integration
# Build ONE pipeline from gm that:
# 1. Keeps rows for years 1998 and 2008 (year %in% c(1998, 2008))
# 2. Drops rows missing life_expectancy (drop_na())
# 3. Groups by continent and year, then summarizes mean_le = mean(life_expectancy)
#    (use .groups = "drop")
# 4. Pivots wider: names_from = year, values_from = mean_le, names_prefix = "y"
# 5. Adds gain = y2008 - y1998 and arranges by gain, largest first
#
# AI Corner: ask your AI assistant the same question, then audit its answer.
# Does it filter the years? Handle NAs? Can you verify every verb it used?

```
