---
title: "Week 1 In-Class Exercise: Welcome & Modern Toolkit"
subtitle: "GPH-GU 2182 · Statistical Programming in R"
author: "YOUR NAME & PARTNER NAME"
date: "September 2, 2026"
format:
  html:
    toc: true
    number-sections: true
    code-tools:
      source: true
      toggle: false
      caption: "Source Code"
---

::: {.callout-note}
### Instructions
Work in pairs during class. 

* 📥 **[Download `week-01-inclass.qmd`](https://raw.githubusercontent.com/gph-2182/gph-2182.github.io/main/weeks/inclass/week-01-inclass.qmd)** to open and edit in RStudio (or use the **`</> Source Code`** button at the top right).
* Render this Quarto document to HTML and push it to your GitHub repository by the end of class for participation credit.
:::

# Cycle 1: R as a Calculator and Environment Setup

## Basic Arithmetic & Calculations

In the code chunk below, compute the Body Mass Index (BMI) for three study participants:

- Participant A: Weight $70\text{ kg}$, Height $1.75\text{ m}$
- Participant B: Weight $85\text{ kg}$, Height $1.80\text{ m}$
- Participant C: Weight $58\text{ kg}$, Height $1.62\text{ m}$

$$\text{BMI} = \frac{\text{weight in kg}}{(\text{height in meters})^2}$$

```{r bmi-calculations}
# Write your calculations below:

```

## Exploring Built-in Functions

Compute the mean of the three BMIs you calculated above. Round the result to 2 decimal places using `round()`.

```{r bmi-summary}
# Write your summary code below:

```

## R Packages and Documentation

1. Load the `r02pro` package (install it in the Console with `install.packages("r02pro")` if you haven't already).
2. Display the first 6 rows of the `ahp` dataset using the `head()` function.
3. Open the help file for the `median()` function using `?median`.

```{r r02pro-check}
# Load library and inspect data:

```

---

# Cycle 2: Reproducible Reporting with Quarto

## Text Formatting & Markdown

In the space below:

- Write a short 2-sentence paragraph describing the purpose of reproducible research in public health.
- Include at least one **bold** phrase and one *italicized* word.
- Add a bulleted list of two advantages of Quarto over traditional copy-pasting into Word.

*(Type your text response here)*

## Code Chunk Customization

Create an R code chunk below with chunk option `#| echo: true` that creates a vector of systolic blood pressure readings: `120, 134, 118, 142, 128` and calculates their standard deviation using `sd()`.

```{r}
#| echo: true
# Write your code here:

```

---

# Cycle 3: Git & GitHub Workflow

Confirm the following steps with your partner:

- [ ] You have initialized Git in your local R project.
- [ ] You have configured your Git identity in RStudio (`git config --global user.name` and `user.email`).
- [ ] You rendered this `.qmd` document to `.html` without any errors.

---

# Cycle 4: AI Corner — Verification Exercise

An AI assistant was asked to write R code to compute the area of a circle with radius $r = 7.5\text{ cm}$.

The AI generated this code:

```{r ai-generated-code}
# AI-generated code:
radius <- 7.5
area <- 3.14 * (radius * 2)
```

### Questions:

1. Is the AI-generated code correct? What error did it make?
2. Correct the code below using R's built-in constant `pi` and proper formula ($\text{Area} = \pi r^2$):

```{r corrected-code}
# Corrected code:

```

---

## Submission Checklist

- [ ] Both your name and your partner's name are in the `author` field at the top.
- [ ] Click **Render** in RStudio and verify the HTML looks clean.
- [ ] Commit and push your changes to GitHub.

